---
title: "LUEM_assignment3"
author: "Lu Yii Wong & Yujin Song"
date: "2024-03-22"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Setup
```{r setup, warning = FALSE, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(caret)
library(pscl)
library(plotROC)
library(pROC)
library(sf)
library(tidyverse)
library(knitr)
library(kableExtra)
library(tigris)
library(viridis)
library(ggplot2)
```

```{r themes}
mapTheme <- theme(plot.title =element_text(size=12),
                  plot.subtitle = element_text(size=8),
                  plot.caption = element_text(size = 6),
                  axis.line=element_blank(),
                  axis.text.x=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks=element_blank(),
                  axis.title.x=element_blank(),
                  axis.title.y=element_blank(),
                  panel.background=element_blank(),
                  panel.border=element_blank(),
                  panel.grid.major=element_line(colour = 'transparent'),
                  panel.grid.minor=element_blank(),
                  legend.direction = "vertical", 
                  legend.position = "right",
                  plot.margin = margin(1, 1, 1, 1, 'cm'),
                  legend.key.height = unit(1, "cm"), legend.key.width = unit(0.2, "cm"))

plotTheme <- theme(
  plot.title =element_text(size=12),
  plot.subtitle = element_text(size=8),
  plot.caption = element_text(size = 6),
  axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 10),
  # Set the entire chart region to blank
  panel.background=element_blank(),
  plot.background=element_blank(),
  #panel.border=element_rect(colour="#F0F0F0"),
  # Format the grid
  panel.grid.major=element_line(colour="#D0D0D0",size=.75),
  axis.ticks=element_blank())

```


```{r load data}
# dependent variable
inundation <- read.csv("https://raw.githubusercontent.com/luyiiw/LUEM/main/calgary_table/zonal_depe.csv")

# independent variables
elevation <- read.csv("https://raw.githubusercontent.com/luyiiw/LUEM/main/calgary_table/zonal_elev.csv")

flow_acc <- read.csv("https://raw.githubusercontent.com/luyiiw/LUEM/main/calgary_table/zonal_fac.csv")

land_use <- read.csv("https://raw.githubusercontent.com/luyiiw/LUEM/main/calgary_table/zonal_lulc.csv")

slope <- read.csv("https://raw.githubusercontent.com/luyiiw/LUEM/main/calgary_table/zonal_slope.csv")

water_body <- read.csv("https://raw.githubusercontent.com/luyiiw/LUEM/main/calgary_table/zonal_water.csv")

```


# Exploratory Analysis
```{r}
# Creating tables to combine exploratory data
flood <- inundation %>%
  select(OBJECTID, RANGE) %>%
  rename(flood = RANGE) %>%
  mutate(flood = factor(flood))

# 1 = flood, 0 = no flood

#1 Elevation
elev_comb <- elevation %>%
  select(OBJECTID, MEAN) %>%
  rename(mean_elev= MEAN)

#2 Slope
slope_comb <- slope %>%
  select(OBJECTID, MAX) %>%
  rename(max_slope= MAX)
# maybe consider mean for model

#3 Water body
water_comb <- water_body %>%
  select(OBJECTID, MIN) %>%
  rename(min_dist_water= MIN)

#4 Flow accumulation
fac_comb <-  flow_acc %>%
  select(OBJECTID, MEAN) %>%
  rename(mean_fac= MEAN)
#changed from max to mean

#5 Land Use/Permeability
# Use max, it's dummy var, as.factor
lu_comb <- land_use %>%
  select(OBJECTID, MAX)%>%
  rename(land_use = MAX) %>%
  mutate(land_use = factor(land_use))
# Urban = 1 , non-urban = 0

# RECLASSIFYING
lu_comb1 <- land_use %>%
  select(OBJECTID, MEAN)%>%
  rename(land_use1 = MEAN) 

lu_comb1 <- lu_comb1 %>%
  mutate(permeability = case_when(
    land_use1 < 0.25 ~ '0',
    land_use1 >= 0.25 ~ '1')) %>%
  select(-land_use1)

# Removing unnecessary data sets
rm(flow_acc, land_use, slope, water_body)
```

```{r}
# Combining the tables (think of another way to do this) consider merge?
flood_comb <- flood %>%
  inner_join(elev_comb, by = "OBJECTID") %>%
  inner_join(slope_comb, by = "OBJECTID") %>%
  inner_join(water_comb, by = "OBJECTID") %>%
  inner_join(fac_comb, by = "OBJECTID")


# flood_comb <- cbind(flood, elev_comb, slope_comb, water_comb, fac_comb) 

#flood_comb2 <- cbind(flood, elev_comb [, -1], slope_comb [, -1], water_comb [, -1], fac_comb [, -1]) %>%
 # rename(mean_elev = elev_comb [, -1])

floodPlotVariables <- 
  flood_comb %>%
  as.data.frame() %>%
  select(flood, mean_elev, max_slope, min_dist_water, mean_fac) %>%
  gather(variable, value, -flood)

```

```{r plot_variables}
# need to check the labels for flood and not flood 
ggplot(floodPlotVariables %>%
         group_by(flood, variable) %>%
         summarize(mean = mean(value))) + 
     geom_bar(aes(as.factor(flood), 
                  mean, 
                  fill=as.factor(flood)),
              stat="identity") + 
     facet_wrap(~variable, scales = "free_y") +
     scale_fill_manual(values = c("dark green", "dark blue"),
                      labels = c("No Flood","Flood"),
                      name = "") +
    labs(x="Flood", y="Value") 


# Plotting dummy variables
# Assuming flood and land_use are already created as factor variables
dummy_comb <- merge(flood, lu_comb, by = "OBJECTID", all = TRUE)

# Plot the bar graph
ggplot(dummy_comb, aes(x = flood, fill = land_use)) +
  geom_bar(position = "dodge") +
  labs(x = "Flood", y = "Count", fill = "Land Use") +
  scale_fill_manual(values = c("dark green", "dark blue"), 
                    labels = c("Non-Urban", "Urban")) 
#change the colors 
#graph shows that urban areas flood more

```


## Model Building 
### Partition training and test sets
```{r}
# Combine all data sets for model building
model_data <- flood %>%
  inner_join(elev_comb, by = "OBJECTID") %>%
  inner_join(slope_comb, by = "OBJECTID") %>%
  inner_join(water_comb, by = "OBJECTID") %>%
  inner_join(fac_comb, by = "OBJECTID") %>%
  inner_join(lu_comb1, by = "OBJECTID") %>%
  mutate(permeability = as.factor(permeability))

#land cover is dummy var and categorical var


# Creating training and test sets
set.seed(3456)
trainIndex <- createDataPartition(model_data$permeability, p = .70,
                                  list = FALSE,
                                  times = 1)

floodTrain <- model_data[ trainIndex,]
floodTest  <- model_data[-trainIndex,]

## work on this to correct the data set output
```


## Making the binomial model
```{r}
floodModel <- glm(flood ~ ., 
                    family="binomial"(link="logit"), data = floodTrain %>%
                                                            as.data.frame() %>%
                                                            select(-OBJECTID))
summary(floodModel)
```

# MAKE CHANGES FROM HERE ON
## Model validation
```{r}
## check from this point, something wrong, observation out put is 1 and 2 instead of 0 and 1
classProbs <- predict(floodModel, floodTest, type="response")


hist(classProbs)

testProbs <- data.frame(obs = as.factor(floodTest$flood),
                        pred = classProbs)

ggplot(testProbs, aes(x = pred, fill=as.factor(obs))) + 
  geom_density() +
  facet_grid(obs ~ .) + 
  xlab("Probability") +
  ylab("Frequency")+
  geom_vline(xintercept = .5) +
  scale_fill_manual(values = c("dark blue", "dark green"),
                      labels = c("Not Flood","Flood"),
                      name = "")+
  plotTheme
```

## Confusion Matrix
```{r}
testProbs$predClass  = ifelse(testProbs$pred > .5 ,1,0)

caret::confusionMatrix(reference = as.factor(testProbs$obs), 
                       data = as.factor(testProbs$predClass), 
                       positive = "1")
```

Predicted = 0, Observed = 0 —> True Negative (No flood correctly predicted 128)

Predicted = 1, Observed = 1 —> True Positive (Floods correctly predicted 42)

Predicted = 1, Observed = 0 —> False Positive (Incorrectly predicted flood when there was no flood 13)

Predicted = 0, Observed = 1 —> False Negative (Incorrectly predicted no flood when there was a flood 20)


## ROC Curve

```{r roc_curve, message = FALSE, warning = FALSE}
# change obs to numeric
# Assuming 'factor_column' is the name of the factor column you want to convert to numeric

testProbs$obs <- as.numeric(as.character(testProbs$obs))

# why is plot not showing up?
ggplot(testProbs, aes(d = obs, m = pred)) + 
  geom_roc(n.cuts = 50, labels = FALSE) + 
  style_roc(theme = theme_grey) +
  geom_abline(slope = 1, intercept = 0, size = 1.5, color = 'grey') 

#area under the curve
auc(testProbs$obs, testProbs$pred)
```

## Cross-validation
```{r}
ctrl <- trainControl(method = "cv", 
                     number = 100, 
                     p = 0.7, 
                     savePredictions = TRUE)

cvFit <- train(as.factor(preserve) ~ .,  data = preserve %>% 
                                                as.data.frame() %>%
                                                select(-geometry, -Id), 
               method="glm", family="binomial",
               trControl = ctrl)

cvFit

ggplot(as.data.frame(cvFit$resample), aes(Accuracy)) + 
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1)) +
  labs(x="Accuracy",
       y="Count")+
  plotTheme
```


## APPLY MODEL TO DENVER
